### 2022-09-29 Thesis/text/diskussion.tex
final v2

The implemented behaviors were a first step to increasing autonomy, but the behavior planner needs more behaviors and fallbacks. Some behaviors that are not implemented in this thesis incorporate the Intel RSS rules for autonomous driving, like keeping appropriate distances, right-of-way rules, and adaptable speed based on occlusions.
Another area of future work is the creation of dedicated behaviors for multi-robot systems. For example the sharing sensor data with other robots to overcome complete sensor failures or intelligent path planning to avoid collisions when robots can access the planned paths of other robots. 
Also, more work is needed to define a robot's safe state or position after triggering an emergency stop. Reaching these minimum risk states after emergencies is then the duty of the behavior planner.

The results must be reproduced on an actual mobile robot. The layout of the behavior tree will stay the same, but further effort is needed to adapt the action nodes of the tree for a real system.

The behaviors must be more deliberative and robust to achieve more independence from human operators. One key question in autonomous driving is the responsibility of a vehicle in a crash because, with higher levels of automation, the human is not needed anymore. However, a machine can not be responsible for its action because it was programmed to do so. Nevertheless, intelligent, robust, and independent decision-making almost always involves machine learning which makes decisions based on training with dedicated datasets and not explicitly written lines of code. Because a system in which the outcome for a high-risk scenario is unclear, machine learning behaviors may not be suited to be used independently. Therefore, an important work area is incorporating machine learning algorithms into a deterministic system, like a behavior tree. A behavior tree could execute and monitor the behaviors to safeguard the indeterministic behaviors and override them if necessary.
%Further research is needed on how behavior trees can manage a system with autonomous machine learning behaviors inside the tree.

>>>>

Firstly, the results must be reproduced on an actual mobile robot. The layout of the behavior tree will stay the same, but further effort is needed to adapt the action nodes of the tree for a real system.

The implemented behaviors were a first step to increasing autonomy, but the behavior planner needs more behaviors and fallbacks. Some crucial behaviors not implemented in this thesis incorporate the Intel RSS rules for autonomous driving, like keeping appropriate distances, right-of-way rules, and adaptable speed based on occlusions. Further, the implemented collision detection needs to be refined to allow collision point detection to improve the whole recovery behavior's robustness. Also, the strategy for finding nearby alternative goals in the path planning behavior needs to be modified or changed to sample more points in all directions around the given goal.

Another area of future work is the creation of dedicated behaviors for multi-robot systems. For example, robots could share sensor data to overcome complete sensor failures or intelligent path planning to avoid collisions when robots can access the planned paths of other robots. 
Also, more work is needed to define a robot's safe state or position after triggering an emergency stop. Reaching these minimum risk states after emergencies is then the duty of the behavior planner.

Intelligent, robust, and independent decision-making almost always involves machine learning (ML) which makes decisions based on training data and not written lines of code. This property makes ML nondeterministic and, therefore, unsuitable for many high-risk and safety-relevant applications. Therefore, an important area of future work is how ML algorithms can be incorporated into a deterministic system like a behavior tree. A behavior tree could execute and monitor the ML behaviors to safeguard and override them if necessary with explicitly programmed but still highly automated behaviors.
%Further research is needed on how behavior trees can manage a system with autonomous machine learning behaviors inside the tree.

### 2022-09-29 Thesis/text/ergebnisse.tex
final v2

The robot is put into specifically designed scenarios to trigger behavior responses from the behavior tree component. The scenario testing is done to evaluate the efficacy of the implemented behaviors in relation to the functional requirements listed in section \ref{sec:Requirements}.

@@

\begin{figure}[ht]
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{images/house_env.png}
		\caption{Apartment Environment in Gazebo}
		\label{fig:house_gazebo}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{images/world_env.png}
\caption{Enclosed Environment in Gazebo}
		\label{fig:world_gazebo}
	\end{subfigure}
	\caption{Gazebo World Environment}
	\label{fig:three graphs}
\end{figure}

@@

\begin{table}[ht]
	\centering
	\caption{Behavior Scenarios}
	\label{tab:behavior_scenarios}
	\renewcommand{\arraystretch}{1.5}
	\resizebox{0.9\textwidth}{!}{%
		\begin{tabular}{| m{0.1\textwidth} | m{0.19\textwidth}| m{0.4\textwidth} | m{0.3\textwidth}|} 
			\hline
			\textbf{Name} & \textbf{Requirements} & \textbf{Description} & \textbf{Success Criteria}\\ 
			\hline	
			NoPath Found$\_$1 & fn$\_$req7 & Path to goal can not be calculated, robot is standing still & 
			Reset system, calculate path to goal, reach goal \\
			\hline
			NoPath Found$\_$2 & fn$\_$req7, fn$\_$req8 & Goal is unreachable, robot is standing still & Alternative goals, close to the original are tested to be reachable \\
			\hline
			Collision$\_$1 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & Robot is standing still and a collision with the robot is caused & Robot can get out of collision state, navigation to goals still working \\
			\hline
			Collision$\_$2 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & The robot is driving and collides with an undetected obstacle (0.25m/s) & 
			Robot can get out of collision state, navigation to goals still working, undetected obstacle gets added to map \\
			\hline
			Battery$\_$1 & fn$\_$req9 & The robot battery runs low & The robot will not drive to a goal that is outside of its reachable range \\
			\hline	
		\end{tabular}
	}
\end{table}

The scenarios execution is done as follows:

For the Lidar, IMU, and Odometry scenarios, the sensor driver was shut down when reaching the desired speed listed in the scenario.

Scenario NoPathFound1 is a scenario in which the lifecycle state gets transitioned into "inactive" and can not execute the planning action anymore when triggered. The BT has to recover from the inability to plan in this scenario.

NoPathFound$\_$2 is simulating the scenario when the planner is active but fails to find a collision-free path to the desired goal. This scenario tests the ability to find alternative goals and navigate to a nearby goal.

In the Collision$\_$1 scenario, an obstacle gets moved so close to the robot that a collision is registered. The forced collision tests the ability to escape a collision and resume navigation afterward. A new navigation goal is published after the behavior tree exits the Collision Fallback Routine.

Collision$\_$2 scenario is testing the same behavior but during active navigation. A flat obstacle is spawned in the path of the robot. The obstacle is too low to be detected by the laserscanner, and a collision occurs. The robot must be able to recover from the crash and keep on navigating towards the goal with an updated global path in this scenario.

The Battery$\_$1 scenario tests if the robot tries to navigate toward a distant goal with a low battery charge. The simulated battery is emptied to a low level, and a new navigation goal is set. If the battery charge runs below zero percent during the navigation, the test fails.

@@

The results in table \ref{tab:results} show an increase in the successful handling of the test scenarios for most test cases. The system supervision component is very reliable and enables the system to recover from an unexpected sensor failure. The more advanced behaviors implemented can improve the autonomous handling of the scenarios. The Battery Scenario was successful every time it was induced. However, the path planning behavior for finding alternative goals was unsuccessful about 50 percent of the time. The test showed that alternative methods for finding new goals would probably have resulted in better performance. The collision behavior during driving worked well, and the map updates are an essential feature for the success of the scenarios.

Nevertheless, there is a lack of intelligence and autonomy when the robot stands still. A collision is forced, but the implemented behavior relies on the same mechanism as during the driving scenario. The missing differentiation led to a test case where the robot reversed into a wall because it was turned 90 degrees by the collision. Therefore, the test failed because a human operator would be needed to correct the robots position. This lack of robustness was not apparent in the driving collision scenario because the robot's force was insufficient to cause a significant shift in orientation when colliding with obstacles.

\begin{table}[ht]
	\centering
	\caption{Results}
	\label{tab:results}
	\renewcommand{\arraystretch}{1.5}
	\resizebox{0.9\textwidth}{!}{%
		\begin{tabular}{| m{0.11\textwidth} | m{0.1\textwidth}| m{0.35\textwidth} | m{0.33\textwidth}|} 
	 	\hline
	 	\textbf{Name} & \textbf{Number of Runs} & \textbf{Percentage Succesful Without Behavior Planning} & \textbf{Percentage Succesful With Behavior Planning}\\ 
	 	\hline
	 	Lidar$\_$1 & 5 & 0 & 100 \\ 
	 	\hline
	 	Lidar$\_$2 & 5 & 0 & 100 \\ 
	 	\hline
	 	Lidar$\_$3 & 5 & 0 & 100 \\ 
	 	\hline
	 	IMU$\_$1 & 5 & 0 & 100 \\ 
	 	\hline
	 	IMU$\_$2 & 5 & 0 & 100 \\ 
	 	\hline
	 	IMU$\_$3 & 5 & 0 & 100 \\ 
	 	\hline
	 	Odom$\_$1 & 5 & 0 & 100 \\ 
	 	\hline
	 	Odom$\_$2 & 5 & 0 & 100 \\ 
	 	\hline
	 	Odom$\_$3 & 5 & 0 & 100 \\ 
	 	\hline
	 	NoPath Found$\_$1 & 20 & 0 & 100 \\ 
	 	\hline
	 	NoPath Found$\_$2 & 20 & 0 & 55 \\ 
	 	\hline
	 	Collision$\_$1 & 20 & 0 & 95 \\ 
	 	\hline
	 	Collision$\_$2 & 20 & 0 & 100 \\ 
	 	\hline
	 	Battery$\_$1 & 10 & 0 & 100 \\ 
	 	\hline
	 	Motor$\_$1 & 10 & 0 & 0 \\ 
	 	\hline
		\end{tabular}
	}
\end{table}

The architectural design choices meet all the non-functional requirements. Implementing the behavior planning and system supervision system eliminates a single point of failure for the system. Even when the Navigation2 system collapses, the robot is able to move to a degree and, more importantly, stop completely (compare table \ref{tab:nonfn_req}, non$\_$fn$\_$req). The behavior tree and algorithms implemented in the behavior tree are all deterministic (non$\_$fn$\_$req3). The robots' behavior goes beyond pure reactivity when navigating through environments, as the advanced behaviors use a dedicated planning phase before the actions are carried out (non$\_$fn$\_$req4).

>>>>

The robot is put into specific scenarios to trigger behavior responses from the behavior tree component. The scenario testing is done to evaluate the efficacy of the implemented behaviors in relation to the functional requirements listed in section \ref{sec:Requirements}.

@@

\begin{figure}[ht]
	\centering
	\begin{subfigure}[b]{0.495\textwidth}
		\centering
		\includegraphics[width=\textwidth]{images/house_env.png}
		\caption{Apartment Environment in Gazebo}
		\label{fig:house_gazebo}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.495\textwidth}
		\centering
		\includegraphics[width=\textwidth]{images/world_env.png}
		\caption{Enclosed Environment in Gazebo}
		\label{fig:world_gazebo}
	\end{subfigure}
	\caption{Gazebo World Environment}
	\label{fig:three graphs}
\end{figure}

@@

\begin{table}[ht]
	\centering
	\caption{Behavior Scenarios}
	\label{tab:behavior_scenarios}
	\renewcommand{\arraystretch}{1.5}
	\resizebox{0.9\textwidth}{!}{%
		\begin{tabular}{| m{0.12\textwidth} | m{0.19\textwidth}| m{0.4\textwidth} | m{0.35\textwidth}|} 
			\hline
			\textbf{Name} & \textbf{Requirements} & \textbf{Description} & \textbf{Success Criteria}\\ 
			\hline	
			NoPath Found$\_$1 & fn$\_$req7 & Path to goal can not be calculated, robot is standing still & 
			Reset system, calculate path to goal, reach goal \\
			\hline
			NoPath Found$\_$2 & fn$\_$req7, fn$\_$req8 & Goal is unreachable, robot is standing still & Alternative goals, close to the original are tested to be reachable \\
			\hline
			Collision$\_$1 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & Robot is standing still and a collision with the robot is caused & Robot can get out of collision state, navigation to goals still working \\
			\hline
			Collision$\_$2 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & The robot is driving and collides with an undetected obstacle (0.25m/s) & 
			Robot can get out of collision state, navigation to goals still working, undetected obstacle gets added to map \\
			\hline
			Battery$\_$1 & fn$\_$req9 & The robot battery runs low & The robot will not drive to a goal that is outside of its reachable range \\
			\hline	
		\end{tabular}
	}
\end{table}

The scenarios were conducted in different ways described in the following paragraphs.

For the lidar, IMU, and odometry scenarios, the sensor driver was shut down when reaching the desired speed listed in the scenario. The location and goal are given as described in the previous chapter. Each sensor was tested multiple times in three different driving situations.

NoPathFound$\_$1 is a scenario in which the lifecycle state of the global planner gets transitioned into "inactive" and cannot execute the planning action anymore when triggered. The BT has to recover from the inability to plan in this scenario. NoPathFound$\_$2 simulates that the planner is active but fails to find a collision-free path to the desired goal. This scenario tests the ability to find alternative goals and navigate to a nearby goal.

In the Collision$\_$1 scenario, an obstacle gets moved so close to the robot that a collision is registered. The forced collision tests the ability to maneuver out of a collision and resume navigation afterward. A new navigation goal is published after the behavior tree exits the Collision Fallback Routine. Collision$\_$2 tests the same behavior but during active navigation. A flat obstacle is spawned in the path of the robot. The obstacle is too low to be detected by the laserscanner, and a collision occurs. The robot must be able to recover from the crash and keep on navigating towards the goal with an updated global path.

The Battery$\_$1 scenario tests if the robot tries to navigate toward a distant goal with a low battery charge. The simulated battery is emptied to a low level, and a new navigation goal is set. The test fails if the battery charge runs below zero percent during the navigation.

@@

The results in table \ref{tab:results} show an increase in the successful handling of the test scenarios for most test cases. The system supervision component is very reliable and enables the system to recover from an unexpected sensor failure. The more advanced behaviors can improve the autonomous handling of the scenarios. The Battery Scenario was successful every time it was induced. However, the path planning behavior for finding alternative goals was unsuccessful about 50 percent of the time.

The collision behavior during driving worked well, and the map updates are an essential feature for the success of the test scenarios. Nevertheless, there is a lack of intelligence and autonomy when the robot stands still, and a collision is forced. The implemented behavior relies on the same mechanism as during the driving scenario. The missing differentiation led to a test case where the robot reversed into a wall because it was turned 90 degrees by the collision. Therefore, the test failed because a human operator would be needed to correct the robot's position. This lack of robustness was not apparent in the driving collision scenario because the robot's force was insufficient to cause a significant shift in orientation when colliding with obstacles.

\begin{table}[ht]
	\centering
	\caption{Results}
	\label{tab:results}
	\renewcommand{\arraystretch}{1.5}
	\resizebox{0.9\textwidth}{!}{%
		\begin{tabular}{| m{0.11\textwidth} | m{0.11\textwidth}| m{0.235\textwidth} | m{0.19\textwidth}|} 
			\hline
			\multirow{2}{*}{\textbf{Name}} & \multirow{2}{0.11\textwidth}{\textbf{Number of Runs}} & \multicolumn{2}{|c|}{\textbf{Percentage Successful}} \\
			\cline{3-4}
			& & \textbf{Without Behavior Planning} & \textbf{With Behavior Planning} \\
			\hline
			Lidar$\_$1 & 5 & 0 & 100 \\ 
			\hline
			Lidar$\_$2 & 5 & 0 & 100 \\ 
			\hline
			Lidar$\_$3 & 5 & 0 & 100 \\ 
			\hline
			IMU$\_$1 & 5 & 0 & 100 \\ 
			\hline
			IMU$\_$2 & 5 & 0 & 100 \\ 
			\hline
			IMU$\_$3 & 5 & 0 & 100 \\ 
			\hline
			Odom$\_$1 & 5 & 0 & 100 \\ 
			\hline
			Odom$\_$2 & 5 & 0 & 100 \\ 
			\hline
			Odom$\_$3 & 5 & 0 & 100 \\ 
			\hline
			NoPath Found$\_$1 & 20 & 0 & 100 \\ 
			\hline
			NoPath Found$\_$2 & 20 & 0 & 55 \\ 
			\hline
			Collision$\_$1 & 20 & 0 & 95 \\ 
			\hline
			Collision$\_$2 & 20 & 0 & 100 \\ 
			\hline
			Battery$\_$1 & 10 & 0 & 100 \\ 
			\hline
		\end{tabular}
	}
\end{table}

The results show that the functional requirements (as defined in \ref{sec:Requirements}) are met in almost all cases. Fn$\_$req1 to fn$\_$req5 are all completely fulfilled. Fn$\_$req6 is only partly fulfilled as the test showed potential shortcomings of the implemented collision behavior. Fn$\_$req7 satisfies the acceptance criteria, but fn$\_$req8 is only partly met by the implemented behavior to find alternative goals. Finally, the battery behavior meets all acceptance criteria, and fn$\_$req9 is fulfilled.

The architectural design choices meet all the non-functional requirements. Implementing the behavior planning and system supervision system eliminates a single point of failure for the system. Even when the Navigation2 system collapses, the robot can move to a degree and, more importantly, stop completely (compare table \ref{tab:nonfn_req}, non$\_$fn$\_$req). The behavior tree and algorithms implemented in the behavior tree are all deterministic (non$\_$fn$\_$req3). The robots' behavior goes beyond pure reactivity when navigating through environments, as the advanced behaviors use a dedicated planning phase before the actions are carried out (non$\_$fn$\_$req4).

### 2022-09-29 Thesis/text/implementierung.tex
final v2

%This is done with the addition of the Decision Gate, that can selectively forward speed commands to the robot controller. The behavior tree's commands have a higher priority and disable the forwarding of the Navigation2 commands. 
%The behavior tree can access the sensor data when needed through the Data Backup component. The System Supervision component controls the health of all components, but the connections are not depicted in the component diagram for reasons of clarity and comprehensibility.

@@

Already existing components are colored in gray, while the components that were implemented in this thesis (i.e., Data Backup, Decision Gate, Behavior Tree and System Supervisor) are depicted in blue. The new components and their functions are described in the following sections of this chapter. The System Supervisor is explained in the BT section as a significant part of the component's functionality is realized inside of the behavior tree.

@@

The data backup component functions as a way to efficiently gather data from all system components and provide necessary data to the behavior tree when the planning process requires past data points. This component subscribes to all sensor data and saves the messages in a queue data structure. The most recent data gets pushed into the array, and if the array exceeds a specified length, the oldest data this data gets deleted. There are two main reasons to limit the array size in this way. Firstly, the goal is to decrease the size of the message that gets sent to the BT to limit the network load. Secondly, the BT often does not need to look back further than a few seconds to make a decision. Subsequently, the computational and network load are decreased significantly. The saved data sources and corresponding lookback times are listed in table \ref{tab:data_backup_types}.

@@

This spinning has to be done in regular intervals. Otherwise, the node might miss incoming messages from subscribed topics. Having this action inside of the behavior tree thread would completely block the execution of the entire autonomy system. If one tries to spin a node only once when a node gets ticked in the BT, losing entire messages is relevant concern because the BT might need longer than expected to return to a node to tick it. That way, the BT node responsible for saving all the relevant data cannot guarantee that it has received all the messages between spins as multiple messages might have been received between the two ticks. The synchronous nature of the ROS executor is why the data backup is not stored within the BT, but in in an external ROS package to seperate the threads.

@@

A simplified behavior tree only depicting the top-level behavior control fallback nodes is shown in figure \ref{fig:top_level_bt}. The tree receives the initial tick signal from the root in an infinite loop.

@@

The system supervisor is using a "Watchdog" design pattern depicted in figure \ref{fig:watchdog_pattern}. A watchdog is used to monitor a system from the outside and take corrective measures when problems are detected.

@@

The system supervisor component is partly outsourced into a component outside of the BT's thread for the same reasons as the sensor backup component discussed in the previous section. The synchronous executor would stop the execution of the whole behavior tree, which is why the component needs to be moved inside of a new thread. The system supervisor, also called the execution checker, is receiving messages from nodes that are not implemented as ROS2 lifecycle nodes. If a node stops sending messages or responding completely, the external execution checker can inform the BT via a provided service of a problem with the node.

ROS2 lifecycle nodes provide information about their health via an FSM implementation inside the node. The current node state can be requested via a service call. However, the node's output needs to be checked, too, because it might be that the FSM inside of the node does not pick up on an error and fails to change the internal lifecycle state.

@@

The system supervision is constantly checking the system's health. The BT is using service calls to get the health information about the components checked inside Condition Nodes. The components get checked by the BT inside of a fallback node. If the condition for the execution is met, the condition node returns "Success", and the fallback would exit to move on to the next fallback and execution condition check as in the lidar fallback depicted in figure  \ref{fig:lidar_fallback} . If the execution checker detects a node failure, a sequence to react to the sensor failure is executed in which the robot slows down to a minimum and tries to restart the sensor. Should the restart fail twice, the robot will not be able to navigate safely anymore and will come to a complete stop. This behavior pattern is realized for the lidar, the IMU, and the wheel odometry.

@@

In order to combat this weakness, the robot must be able to detect that a collision occurred, get out of the collision state, update the map and reset all costmaps to generate a new path to the original goal.

@@

The activity of the "MapUpdateAfterCollision" action node is depicted in figure \ref{fig:activity_diagram_collision}. The collision point is estimated by the robot's pose, which refers to the robot base link and the robot footprint. It is assumed that the collision point was right in front of the robot because the local planner favors forward drive kinematics and probably is driving forwards when a collision occurs. The simulated bumper sensor can provide more nuanced information about the point of contact, but this functionality can only be achieved with high developmental effort on a real robot with an IMU-based collision detection. In order to limit the effort of transfering and running the implemented software on a real robot, the obstacle position is simplified to always be in front of the robot. A $25$ by $25$ $cm$ square with $100$ percent occupation at the collision point is added to the current map and then saved to be reloaded by the map server. The original goal is then republished to replan the global path on the updated map. A more detailed view of the whole behavior is depicted in figure \ref{fig:activity_diagram_collision}.

%The collision position is calculated with 
%
%\textit{Formula here}

%as depicted in figure xx. 
% add the 
\begin{figure*}[ht!]
	\centering
	\includegraphics[width=0.90\textwidth]{images/activity_diagram_collision.png}
	\caption{Activity Diagram for the Collision Behavior}
	\label{fig:activity_diagram_collision}
\end{figure*}

This way, obstacles lower than the lidar scan height can be detected. With this behavior, the robot should be able to eventually navigate around transparent obstacles that do not reflect the laser. Glass doors are known to cause issues for lidar-based robots, because of their reflective properties (see section)

@@

The battery behavior aims to monitor the robot's battery and has to decide if the robot should navigate to a new goal location or not. The decision is based on the current battery charge and the estimated consumption of energy to reach the new goal. The robot should not begin the navigation process if it is probable that the remaining charge will not suffice to arrive at the goal. Figure \ref{fig:battery_fallback} depicts the behavior tree fallback.

@@

For the simulation, a battery package was created, which updates the battery charge every second. The battery package subscribes to the command velocity topic and decreases the battery charge for every message received. The battery package was created because the behavior needs to be tested and Gazebo does not offer this functionality. The consumption rates are not close to reality but are chosen to allow for a rapid testing of the behavior. The average values for idle and driving consumption on a real robot could be experimentally determined.

@@

With this behavior, a way to make the global planner more flexible was implemented. The execution checker component monitors the health of the Nav2 global planner. Additionally, the behavior tree analyzes the global planner's output. If the latter condition fails, the given navigation goal is not reachable by the planning algorithm. By shifting the goal to a nearby alternative goal, the robot tries to navigate to close location, although the original goal cannot be reached.

If the location is still unreachable, the goal gets moved further away from the original. The routine repeats up to ten times, leading to a maximum distance to the original goal of up to one meter. The goal gets moved on a straight vector towards the robot. Other strategies for finding alternative goals were not implemented in this behavior due to time-constraints. Figure \ref{fig:global_planning_fallback} depicts the behavior tree fallback for the behavior.

@@

For testing purposes, the simulated sensor data is not published directly on the actual topic by Gazebo. An intermediary node is used, which subscribes to the Gazebo topic and republishes the information on the original topic. The failure of a sensor can be simulated this way by stopping the execution of the intermediary node. Figure \ref{fig:rqt_graph} depicts the ROS nodes and their subscribed and published topics. Topics are depicted as rectangles and nodes are shown as circles.

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.7\textwidth]{images/rqt_graph_sensor_drivers.png}
	\caption{Graph for the Sensor Driver Nodes and Topics}
	\label{fig:rqt_graph}
\end{figure}

@@

Because of the necessity to have ROS nodes spinning outside of the BT thread, as discussed in section \ref{subsec:BT System Supervision}, additional interfaces are created in order to receive information through service calls. The created service types are listed in table \ref{tab:custom_interfaces}. All service definitions are created in a dedicated package named "bt$\_$msgs".

>>>>

%This is done with the addition of the Decision Gate, that can selectively forward speed commands to the robot controller. The behavior tree's commands have a higher priority and disable the forwarding of the Navigation2 commands. 
%The behavior tree can access the sensor data through the Data Backup component. The System Supervision component controls the health of all components, but the connections are not depicted in the component diagram for clarity and comprehensibility.

@@

Already existing components are colored in gray, while the components implemented in this thesis (i.e., Data Backup, Decision Gate, Behavior Tree, and System Supervisor) are depicted in blue. The new components and their functions are described in the following sections of this chapter. The System Supervisor is explained in the BT section as a significant part of the component's functionality is realized inside the behavior tree.

@@

The data backup component functions as a method to efficiently gather data from all system components and provide it to the behavior tree when the planning process requires past data points. This component subscribes to all sensor data and saves the messages in a queue data structure. The most recent data gets pushed into the array, and if the array exceeds a specified length, the oldest data this data gets deleted. There are two main reasons to limit the array size in this way. Firstly, the goal is to decrease the size of the message that gets sent to the BT to limit the network load. Secondly, the BT often does not need to look back further than a few seconds to make a decision. Subsequently, the computational and network load are decreased significantly. The saved data sources and corresponding lookback times are listed in table \ref{tab:data_backup_types}.

@@

This spinning has to be done in regular intervals. Otherwise, the node might miss incoming messages from subscribed topics. This action inside the behavior tree thread would completely block the execution of the entire autonomy system. If one tries to spin a node only once when a node gets ticked in the BT, losing entire messages is a relevant concern because the BT might need longer than expected to return to a node to tick it. That way, the BT node responsible for saving all the relevant data cannot guarantee that it has received all the messages between spins as multiple messages might have been received between the two ticks. The synchronous nature of the ROS executor is why the data backup is not stored within the BT, but in an external ROS package to separate the threads.

@@

A simplified behavior tree only depicting the top-level behavior control fallback nodes is shown in figure \ref{fig:top_level_bt} on page \pageref{fig:top_level_bt}. The tree receives the initial tick signal from the root in an infinite loop.

@@

The system supervisor uses a "Watchdog" design pattern depicted in figure \ref{fig:watchdog_pattern}. A watchdog is used to monitor a system from the outside and take corrective measures when problems are detected.

@@

The system supervisor component is partly outsourced into a component outside of the BT's thread for the same reasons as the sensor backup component discussed in the previous section. The synchronous executor would stop the execution of the whole behavior tree, which is why the component needs to be moved inside of a new thread. The system supervisor, also called the execution checker, receives messages from nodes not implemented as ROS2 lifecycle nodes. If a node stops sending messages or responding completely, the external execution checker can inform the BT via a provided service of a problem with the node.

ROS2 lifecycle nodes provide information about their health via an FSM implementation inside the node. The current node state can be requested via a service call. However, the node's output needs to be checked, too, because it might be that the FSM inside the node does not pick up on an error and fails to change the internal lifecycle state.

@@

The system supervision is constantly checking the system's health. The BT is using service calls to get the health information about the components checked inside Condition Nodes. The components get checked by the BT inside of a fallback node. If the condition for the execution is met, the condition node returns "Success", and the fallback would exit to move on to the next fallback and execution condition check as in the lidar fallback depicted in figure  \ref{fig:lidar_fallback}. If the execution checker detects a node failure, a sequence to react to the sensor failure is executed in which the robot slows down to a minimum and tries to restart the sensor. Should the restart fail twice, the robot will not be able to navigate safely anymore and will come to a complete stop. This behavior pattern is realized for the lidar, the IMU, and the wheel odometry.

@@

In order to combat this weakness, the robot must be able to detect that a collision occurred, get out of the collision state, update the map and reset all costmaps to generate a new path to the original goal. The described behavior is depicted in figure \ref{fig:collision_fallback}.

@@

The activity of the "MapUpdateAfterCollision" action node is depicted in figure \ref{fig:activity_diagram_collision} on page \pageref{fig:activity_diagram_collision}. The collision point is estimated by the robot's pose, which refers to the robot base link and the robot footprint. It is assumed that the collision point was right in front of the robot because the local planner favors forward drive kinematics and probably is driving forwards when a collision occurs. The simulated bumper sensor can provide more nuanced information about the point of contact. However, this functionality can only be achieved with high developmental effort on a real robot with IMU-based collision detection. In order to limit the effort of transferring and running the implemented software on a real robot, the obstacle position is simplified to always be directly in front of the robot. A $25$ by $25$ $cm$ square with $100$ percent occupation at the collision point is added to the current map and then saved to be reloaded by the map server. The original goal is then republished to replan the global path on the updated map..

%The collision position is calculated with 
%
%\textit{Formula here}
%as depicted in figure xx. 
% add the 
\begin{figure*}[ht!]
	\centering
	\includegraphics[width=0.90\textwidth]{images/activity_diagram_collision.png}
	\caption{Activity Diagram for the Collision Behavior}
	\label{fig:activity_diagram_collision}
\end{figure*}

This way, obstacles lower than the lidar scan height can be detected. With this behavior, the robot should be able to eventually navigate around transparent obstacles that do not reflect the laser. Glass doors are known to cause issues for lidar-based robots because of their reflective properties (see section \ref{subsec:nav2}).

@@

The battery behavior aims to monitor the robot's battery and has to decide if the robot should navigate to a new goal location or not. The decision is based on the current battery charge and the estimated consumption of energy to reach the new goal. The robot should not begin the navigation process if it is probable that the remaining charge will not suffice to arrive at the goal. Figure \ref{fig:battery_fallback} on page \pageref{fig:battery_fallback} depicts the battery fallback.

@@

A battery package was created for the simulation, which updates the battery charge every second. The battery package subscribes to the command velocity topic and decreases the battery charge for every message received. The battery package was created because the behavior needs to be tested, but Gazebo does not offer this functionality. The consumption rates are not close to reality but are chosen to allow for rapid testing of the behavior. The average values for idle and driving consumption on a real robot could be experimentally determined.

@@

With this behavior, a way to make the global planner more flexible was implemented. The execution checker component monitors the health of the Nav2 global planner. Additionally, the behavior tree analyzes the global planner's output. If the latter condition fails, the given navigation goal is not reachable by the planning algorithm. By shifting the goal to a nearby alternative goal, the robot tries to navigate to a close location, although the original goal cannot be reached.

If the location is still unreachable, the goal gets moved further away from the original. The routine repeats up to ten times, leading to a maximum distance to the original goal of up to one meter. The goal gets moved on a straight vector towards the robot. This behavior did not implement other strategies for finding alternative goals due to time constraints of the thesis. Figure \ref{fig:global_planning_fallback} depicts the behavior tree fallback for the behavior.

@@

For testing purposes, the simulated sensor data is not published directly on the actual topic by Gazebo. An intermediary node is used, which subscribes to the Gazebo topic and republishes the information on the original topic. The failure of a sensor can be simulated this way by stopping the execution of the intermediary node. 
%Figure \ref{fig:rqt_graph} depicts the ROS sensor nodes and their subscribed and published topics. Topics are illustrated as rectangles and nodes are shown as circles. The "/clock" and "/rosout" topic are unrelated to the described functionality and are required to exist to work with the simulation and ROS.
%
%\begin{figure}[ht]
%	\centering
%	\includegraphics[width=0.7\textwidth]{images/rqt_graph_sensor_drivers.png}
%	\caption{Graph for the Sensor Driver Nodes and Topics}
%	\label{fig:rqt_graph}
%\end{figure}

@@

Because of the necessity to have ROS nodes spinning outside of the BT thread, as discussed in section \ref{subsec:BT System Supervision}, additional interfaces are created to receive information through service calls. The created service types are listed in table \ref{tab:custom_interfaces}. All service definitions are created in a dedicated package named "bt$\_$msgs".
